---
title: "Assignment 2"
author: "Steve Hawley"
date: "November 18, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section 0: Load the data and packages

```{r get_data, message=FALSE, results='markup', warning=FALSE}
library(foreign) #for importing spss data
library(summarytools) #for descriptive and cross tables
library(dplyr) #for grouping and plotting
library(ggplot2) #for plotting
library(effsize) #for cohen's d
library(car) #for leven's test
library(sjstats) #for eta squared
library(tidyr) #for gather function

# import the spss file. Label the factors
data<-read.spss("Data+for+assignments.sav",to.data.frame = TRUE, add.undeclared.levels = "no")

#subset data for analysis
rma<-data %>% 
  select(code,age,gender,satcurwt,satwt18) %>% 
  filter(!is.na(satcurwt),!is.na(satwt18),!is.na(gender), age>18) %>% #filtering out missing values and age<=18 (since one of the questions requires respondants to be at least 18)
  mutate(ID = seq.int(n())) #have to re-assign unique IDs since the code variable is not unique for some reason

# take a look at the data. See what the fields are
head(rma)

#inspect the data types
str(rma)


```

## Section 1: Pair-Samples t-test

#### **Research Question:** Is there a significant change in weight satisfaction in respondants' current weight compared to their weight at 18 years old?  


```{r paired_samples, results='asis'}

#subset the variables for paired samples t-test
pst<-rma %>% select(satcurwt,satwt18)

#get the descriptives
descr(pst, style = 'rmarkdown', transpose = TRUE)

```
```{r paired_samples2}
#run the t test
t.test(pst$satwt18, pst$satcurwt, paired = TRUE)

#effect size
cohen.d(pst$satwt18, pst$satcurwt, paired = TRUE)

```
####Results (APA Style)
A paired-samples _t_ test was conducted to evaluate whether the respondants experienced a change in life satisfaction between 18 years old and their current age. The results indicated that the mean weight satisfaction at age 18 ( _M_=7.09, _SD_=2.57) was significantly greater than the mean weight satisfaction at the respondants' current age ( _M_=5.94, _SD_=2.70), _t_(647) = 10.233, _p_<0.001. The standarized effect size index, _d_, was 0.40, which indicates a small effect size of this difference. The 95% confidence interval for the mean difference between the weight satisfaction at 18 and current age was 0.29 to 0.51.

## Section 2: Repeated-Measures ANOVA

#### **Research Question:** Is there a significant change in weight satisfaction in respondants' current weight compared to their weight at 18 years old?  


```{r repmea_aov}

#Following the approach for repeated measures anova found here: 
#http://www.cookbook-r.com/Statistical_analysis/ANOVA/#anovas-with-within-subjects-variables
#First I need to reshape the data from wide to long for this to work.
#This is where I'll need the unique ID variable I created earlier. I'll also need to put satwt18 and satcurwt into a single column as labels, with 'value' as their values.
longdata<-gather(rma,satis,value,satwt18:satcurwt)

#let's take a look
head(longdata)

#set up the data to plot. Order the factors
longdata$satis<-factor(longdata$satis, levels = c("satwt18","satcurwt"))

#group the data for plotting
p<-longdata %>% group_by(satis) %>% summarise(satscore=mean(value))

#line plot to look at the change in weight satisfaction between age 18 and current weight
p %>%
  ggplot(aes(x=satis, y=satscore, group=1)) +
  geom_line() +
  geom_point() +
  expand_limits(y=c(0,10)) +
  xlab("")

#need to set the two timepoints (i.e., repeated IDs from the previously wide variables) as factors
longdata$ID<-factor(longdata$ID)

#Run the ANOVA
aov_long<-aov(value~satis + Error(ID/satis), data=longdata)
summary(aov_long)

eta_sq(aov_long)

```
####Results (APA Style)
A one-way within-subjects ANOVA was conducted with the factor being age and the dependant variable being weight satisfaction score. The results of ANOVA indicated a significant age effect, F(1,647) = 104.7 , p<0.001, multivariate eta-squared = 0.045. No pairwise comparisons were conducted, as only two groups were present in the ANOVA. Thus, the ANOVA shows that the mean weight satisfaction at 18 years old was significantly greater ( _M_=7.09, _SD_=2.57) than the mean weight satisfaction at the respondants' current age ( _M_=5.94, _SD_=2.70).

## Section 3: Mixed ANOVA

#### **Research Question:** Is there a significant change in weight satisfaction in respondants' current weight compared to their weight at 18 years old and is this change different for males and females?


```{r mixed_aov}

#For the mixed anova, I'll follow the approach from the same site as the previous section: 
#http://www.cookbook-r.com/Statistical_analysis/ANOVA/#mixed-design-anova
#Run the anova, this time with an additional factor

#break down the descriptives to male and female
pst_m<-rma %>% filter(gender=='male')%>% select(satcurwt,satwt18) 
pst_f<-rma %>% filter(gender=='female') %>% select(satcurwt,satwt18) 

descr(pst_m, style = 'rmarkdown', transpose = TRUE)
descr(pst_f, style = 'rmarkdown', transpose = TRUE)


#group the data for plotting
p<-longdata %>% group_by(satis, gender) %>% summarise(satscore=mean(value))

#let's see if the lines are (roughly) parellel (i.e., no interaction effect)
p %>%
  ggplot(aes(x=satis, y=satscore, color=gender, group=gender)) +
  geom_line() +
  geom_point() +
  expand_limits(y=c(0,10)) +
  xlab("") +
  scale_y_continuous(breaks = c(0,2,4,6,8,10))

aov.mix<-aov(value~satis*gender + Error(ID/(satis*gender)), data=longdata)
summary(aov.mix)

eta_sq(aov.mix)

```
####Results (APA Style)
A mixed-design ANOVA with age as a within-subjects effect and gender as a between-subjects effect was conducted to evalute the change in weight satisfaction between age 18 and current age between males and females. The results of then ANOVA indicated no significant interaction between age and gender F(1,646) = 0.57, p = 0.45, multivariate eta_squared = 0.00. The results of ANOVA indicated a significant age effect, F(1,647) = 104.7 , p<0.001, multivariate eta-squared = 0.045. and a significant gender effect F(1,646) = 12.69, P<0.001, multivariate eta-squared = 0.013. No pairwise comparisons were conducted, as only two groups were present for each effect in the ANOVA. Thus, the ANOVA shows that the mean weight satisfaction at 18 years old was significantly greater than the mean weight satisfaction at the respondants' current age for both genders, with males having significantly greater weight satisfaction scores at both time points.